---
title: "prova 2"
output: html_notebook
---

Pergunta 1
Suppose I define the following function in R


cube <- function(x, n) {
        x^3
}

What is the result of running

cube(3)


```{r}

cube <- function(x, n) {
        x^3}
cube(3)
```
Pergunta 2
The following code will produce a warning in R.
Why?


```{r}
x <- 1:10
if(x > 5) {
        x <- 0
}
```
Pergunta 3
Considere a seguinte função

```{r}
f <- function(x) {
        g <- function(y) {
                y + z
        }
        z <- 4
        x + g(x)
}
```

Se eu correr em R
o que retorna 
```{r}
z <- 10
f(3)
```
Pergunta 4
Considere a seguinte expressão:
Qual é o valor de 'y' após avaliar esta expressão?
```{r}
x <- 5
y <- if(x < 3) {
        NA
} else {
        10
}
y
```
Pergunta 5
Consider the following R function
Which symbol in the above function is a free variable?

```{r}
h <- function(x, y = NULL, d = 3L) {
        z <- cbind(x, d)
        if(!is.null(y))
                z <- z + y
        else
                z <- z + f
        g <- x + y / z
        if(d == 3L)
                return(g)
        g <- g + 10
        g
}


z
g

```

Pergunta 6
What is an environment in R?
R: a collection of symbol/value pairs



Pergunta 7
The R language uses what type of scoping rule for resolving free variables?
R: lexical scoping


Pergunta 8
How are free variables in R functions resolved?
R: The values of free variables are searched for in the environment in which the function was defined


Pergunta 9
What is one of the consequences of the scoping rules used in R?
R:All objects must be stored in memory



Pergunta 10
In R, what is the parent frame?
R:It is the environment in which a function was called

