---
title: 'Part 2: Basic Inferential Data Analysis Instructions'
output:
  html_document:
    keep_md: yes
  pdf_document: default
---

# OVERVIEW

In the second portion of the project, we're going to analyze the ToothGrowth data in the R datasets package

```{r setup, include = FALSE}
library(ggplot2)
library(datasets)

knitr::opts_chunk$set(echo = TRUE, fig.path='figures/')


# disable scientific notation for numbers
options(scipen = 1)
```


# DATA SUMMARY

After loading the ToothGrowth dataset, provide a basic summary of the data.

```{r basic-data-summary, echo = TRUE}

# load the sample dataset containing ToothGrowth data
data ("ToothGrowth")

str(ToothGrowth)

summary(ToothGrowth)

```
```{r}
head(ToothGrowth, 10)
```


```{r}
# tabulate delivery method and dosage level values
table(ToothGrowth$supp, ToothGrowth$dose)
```


```{r}
# summary of tooth length data grouped by delivery method and dosage level
by(data = ToothGrowth$len, INDICES = list(ToothGrowth$supp, ToothGrowth$dose), summary)
```


```{r}
```



# EXPLORATORY DATA ANALYSIS

Perform some basic exploratory data analyses of the data. The analyses will
explore the following relations:

1. Tooth Length as a function of Delivery Method
2. Tooth Length as a function of Dosage Level
3. Tooth Length as a function of Delivery Method (supp) and Dosage Level

###  1- Tooth Length to Delivery Method

```{r, echo = TRUE}


levels(ToothGrowth$supp) <- c("Orange Juice", "Ascorbic Acid")

ggplot(data = ToothGrowth, aes(x = supp, y = len)) + 
    geom_boxplot(aes(fill = supp)) +
    xlab("Delivery Method") +
    ylab("Tooth Length") +
    theme(plot.title = element_text(size = 15, hjust = 0.5)) +
    ggtitle("Tooth Length as a Function of Delivery Method") 


```
### Findings

Based on the graph, it is observed that, regardless of the dosage level, the use of orange juice as a method of administration had a more favorable effect on tooth growth than ascorbic acid.


## 2- Tooth Length to Dosage Level

```{r, echo = TRUE}

ggplot(data = ToothGrowth, aes(x = factor(dose), y = len)) + 
    geom_boxplot(aes(fill = factor(dose))) +
    xlab("Dosage Level (mg/day)") +
    ylab("Tooth Length") +
    guides(fill=guide_legend(title="dose")) +
    theme(plot.title = element_text(size = 15, hjust = 0.5)) +
    ggtitle("Tooth Length as a Function of Dosage Level")

```

### Findings

Based on the graph, it is observed that higher levels of vitamin C dosage, regardless of the method of administration, had a more favorable effect on tooth growth.



## 3- Tooth Length to Delivery Method and Dosage Level

```{r, echo = TRUE}


levels(ToothGrowth$supp) <- c("Orange Juice", "Ascorbic Acid")

ggplot(data = ToothGrowth, aes(x = supp, y = len)) + 
    geom_boxplot(aes(fill = supp)) +
    facet_wrap(~ dose) +
    xlab("Delivery Method") +
    ylab("Tooth Length") +
    guides(fill=guide_legend(title="supp")) +
    theme(plot.title = element_text(size = 14, hjust = 0.5, vjust = 0.5),
          axis.text.x = element_text(angle = 45,
                                     hjust = 0.5,
                                     vjust = 0.5,
                                     margin = margin(b = 10))) +
    ggtitle("Tooth Length as a Function of\nDelivery Method and Dosage Level")

```

### Findings

Based on the graph, for dosages of 0.5 to 1 mg/day, orange juice was better. As for the 2mg/day dosage by the median, there was no difference between the means of administration, however if we look at the quartiles, orange juice has a more robust result.



## INFERENTIAL STATISTICS


Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. For this a t-test will be performed 
crescimento do dente.

### Hypothesis 1

Test if whether is any correlation between tooth grow and dogase level  
The null hypothesis that the two delivery methods have no effect on
tooth growth.

```{r, echo = TRUE}

t.test(len ~ supp, data = ToothGrowth, conf.level = 0.95)

```
P-value is 0.06063 (>0.05)
Confidence Interval is -0.17 to 7.57 (contains 0)
This indicates weak evidence against the
null hypothesis so we fail to reject the null hypothesis. So the dosage seems to have no impact on Tooth growth


### Hypothesis 2

Study the impact of delivery method on tooth growth for a single dosage level.
A t-test will be performed by comparing Tooth Grow with Dose Amount looking at the different pairs of dose values

```{r, echo = TRUE}

# Dose amounts 0.5 and 1.0
t.test(len~dose,data=subset(ToothGrowth, ToothGrowth$dose %in% c(1.0,0.5)))

```
P-value is 1.268e-07 (>0.05)
Confidence Interval is -11.983781 to -6.276219 (doesn’t contains 0)
Based on these results, it was found that between 0.05mg/day and 1 mg/day there is a difference in tooth growth



```{r, echo = TRUE}

# Dose amounts 0.5 and 2.0
t.test(len~dose,data=subset(ToothGrowth, ToothGrowth$dose %in% c(2.0,0.5)))
```
P-value is 4.398e-14 (>0.05)
Confidence Interval is -18.15617 to -12.83383 (doesn’t contains 0)
Based on these results, it was found that between 0.05mg/day and 2mg/day there is a difference in tooth growth


```{r, echo = TRUE}

# Dose amounts 1.0 and 2.0
t.test(len~dose,data=subset(ToothGrowth, ToothGrowth$dose %in% c(2.0,1.0)))

```
P-value is 0.00001906 (>0.05)
Confidence Interval is -8.996481 to -3.733519 (doesn’tcontains 0)
Based on these results, it was found that between 0.05mg/day and 1 mg/day there is a difference in tooth growth


# CONCLUSION

## Assumptions

 - Tooth growth follows a normal distribution.

 - The sample is representative of the population


## Conclusions

Based on exploratory data analysis and confirmed by hypothesis tests and associated confidence intervals, we can safely infer that an increase in dosage levels of vitamin C increases tooth growth. 
However, the delivery method had any effect on tooth growth.
